\chapter*{Introduction}
%% Unlike \chapter, \chapter* does not update the headings and does not
%% enter the chapter into the table of contents. If we want correct
%% headings and a table of contents entry, we must add them manually:
\markright{\textsc{Introduction}}
\addcontentsline{toc}{chapter}{Introduction}

There is a long history of neuroscientific studies viewing the nervous system as a complex network \cite{sporns_structure_2013}. This thesis aims to explore methodologies rooted in complex network analysis, specifically network communication models, to study both empirical and simulated electroencephalography (EEG) recordings of brain responses to transcranial magnetic stimulation (TMS). 

The network communication models aim to provide a high-level explanation of the information transfer in the brain. They build a connection between a static structural connectivity of the brain and its function, represented here by spatiotemporally variable response to a stimulus. 

We chose the recent paper \textit{Communication dynamics in the human connectome shape the cortex-wide propagation of direct electrical stimulation} by Seguin et al. \cite{seguin_communication_2023} as an entry point to this topic. The paper presents results capturing some of the relationships between structural connectivity and stimulus response based on network communication models. The work, done on invasive intracranial data in a large cohort, serves as a guide for the application of the network communication models methodology to noninvasive data. 

Our aim here is to investigate to which degree these results can be translated to the noninvasive setting of TMS-EEG. We propose several possibilities for the characterization of the TMS-evoked response, which are later correlated with the network communication models. Besides empirical TMS-EEG measurements, we also examine their simulated equivalent created by Momi et al. \cite{momi_tms-evoked_2023}. We compare the results for empirical and synthetic data. 

The thesis is organized as follows. Chapter \ref{ch:brain} provides the necessary information about the biology of the human brain and technical aspects of the measuring techniques employed in data acquisition. This is immediately followed by a description of the network construction from the measurements of the brain in Chapter \ref{ch:SC}.

Chapter \ref{ch:networks} is devoted to the description of the network communication models. It includes their mathematical formulations, and it relates them to the biological aspects of communication processes in the brain.

Chapter \ref{ch:ftract} includes a replication of the results by Seguin et al. and our extension of their work. They focused on whole-brain stimulus-response probabilities, but we also examined the case considering only one stimulated region. 

The next chapters are dedicated to the application of the communication network methodology to both empirical and simulated noninvasive EEG recordings of brain responses to transcranial magnetic stimulation and the comparison of obtained results with the original work on intracranial data. Chapter \ref{ch:pytepfit} discusses several possibilities of TMS-EEG data description. It shows the results of the application of the approach proposed by Seguin et al. to these data characteristics. It also compares the results for empirical and simulated TMS-evoked potentials. Followingly, Chapter \ref{ch:compare} compares the results for intracranial and transcranial data.

Last but not least, Chapter \ref{ch:SC_indepth} focuses on a robustness analysis of the results, especially regarding variations in structural connectivity datasets and several methods of group-averaging structural connectivity data for the datasets where individual subject matrices are available.


